Redshift evolution of the underlying type Ia supernova stretch distribution

### 2021-03-16 Nicolas.tex
80 char limit in .tex
Wrong date in Rigault 2020
Removed bold

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties arises, notably that of the evolution of SN~Ia populations. We
    study the dependence on redshift of the SN~Ia \texttt{SALT2.4} light-curve
    stretch, which is a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to be strongly correlated with
    the SN environment, notably with stellar age tracers. We modeled the
    underlying stretch distribution as a function of redshift, using the
    evolution of the fraction of young and old SNe~Ia as predicted using the
    SNfactory dataset, and assuming a constant underlying stretch distribution
    for each age population consisting of Gaussian mixtures. We tested our
    prediction against published samples that were cut to have marginal
    magnitude selection effects, so that any observed change is indeed
    astrophysical and not observational in origin. In this first study, there
    are indications that the underlying SN~Ia stretch distribution evolves as a
    function of redshift, and that the \textbf{age} drifting model is a better
    description of the data than any time-constant model, including the
    sample-based asymmetric distributions that are often used to correct
    Malmquist bias at a significance higher than 5 $\sigma$. The favored
    underlying stretch model is a bimodal one, composed of a high-stretch mode
    shared by both young and old environments, and a low-stretch mode that is
    exclusive to old environments. The precise effect of the redshift evolution
    of the intrinsic properties of a SN Ia population on cosmology remains to be
    studied. The astrophysical drift of the SN stretch distribution does affect
    current Malmquist bias corrections, however, and thereby the distances that
    are derived based on SN that are affected by observational selection
effects. We highlight that this bias will increase with surveys covering
increasingly larger redshift ranges, which is particularly important for the
Large Synoptic Survey Telescope.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
enabled the discovery that the Universe\textbf{'s expansion accelerates}
\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today
for understanding the properties of dark energy (DE) because they are the only
tool that can precisely map the recent expansion rate ($z<0.5$) when DE is
driving it \citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the
key to directly measuring the Hubble constant ($H_0$) if their absolute
magnitude can be calibrated \citep{riess2016, freedman2019}. Interestingly, the
value of $H_0$ that is derived when the SNe~Ia are anchored to Cepheids
\citep[the Supernovae, $H_0$, for the Equation of State of dark energy
project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
from cosmic microwave background (CMB) data measured by Planck assuming the
standard $\Lambda$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
luminosity is anchored at intermediate redshift by the baryon acoustic
oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
branch technique in place of the Cepheids seems to favor an intermediate value
of $H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
lensing also appear to favor high $H_0$ values \citep{wong2019}.

@@

To model the evolution of the full SN stretch distribution as a function of
redshift, we need to model the SN stretch distribution for each age subsample
given our aforementioned model of the evolution of the fraction of younger and
older SNe~Ia with cosmic time. \cite{rigault2020} presented the relation between
SN stretch and LsSFR measurement, a progenitor age tracer, using the SNfactory
sample. This relation is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory
SNe used in the current analysis. Given the structure of the stretch-LsSFR
scatter plot, our model of the underlying SN~Ia stretch distribution is defined
as follows: \textbf{the stretch distribution of the younger population
($\log(\mathrm{LsSFR})\geq-10.82$) is modeled as a single normal distribution
$\mathcal{N}(\mu_1, \sigma_1{}^2)$, and the stretch distribution of the older
population ($\log(\mathrm{LsSFR})<-10.82$) is modeled as a bimodal Gaussian
mixture $a\times \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2{}^2)$, where one mode is the same as for the young
population, $a$ representing the relative effect of the two modes.}

@@

\begin{table*}
    \centering
    \caption{Best-fit values of the parameters for the base stretch distribution
    model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
569 SN~Ia sample, or the conservative \textbf{sample} (422).}
    \label{tab:modelresults}
    \begin{tabular}{lccccc}
        \hline\hline
        Sample & $\mu_1$ & $\sigma_1$
               & $\mu_2$ & $\sigma_2$
               & $a$ \\
        \hline
        SNfactory & $ 0.41 \pm 0.08$ & $0.55 \pm 0.06$
                  & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$
                  & $ 0.48 \pm 0.08$ \\
        Fiducial & $ 0.37 \pm 0.05$ & $0.61 \pm 0.04$
                 & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$
                 & $ 0.51 \pm 0.09$ \\
        Conservative & $ 0.38 \pm 0.05$ & $0.60 \pm 0.04$
                     & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$
                     & $ 0.47 \pm 0.09$ \\
        \hline
    \end{tabular}
\end{table*}

@@

%\bibitem[Poulin et al.(2019)]{poulin2019} Poulin, V., Smith, T.~L., Karwal, T., et al.\ 2019, \prl, 122, 221301

@@

\bibitem[Rigault et al.(2020)]{rigault2020} Rigault, M., Brinnel, V., Aldering,
G., et al.\ 2029, \aap, 644, A176

>>>>

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties arises, notably that of the evolution of SN~Ia populations. We
    study the dependence on redshift of the SN~Ia \texttt{SALT2.4} light-curve
    stretch, which is a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to be strongly correlated with
    the SN environment, notably with stellar age tracers. We modeled the
    underlying stretch distribution as a function of redshift, using the
    evolution of the fraction of young and old SNe~Ia as predicted using the
    SNfactory dataset, and assuming a constant underlying stretch distribution
    for each age population consisting of Gaussian mixtures. We tested our
    prediction against published samples that were cut to have marginal
    magnitude selection effects, so that any observed change is indeed
    astrophysical and not observational in origin. In this first study, there
    are indications that the underlying SN~Ia stretch distribution evolves as a
    function of redshift, and that the age drifting model is a better
    description of the data than any time-constant model, including the
    sample-based asymmetric distributions that are often used to correct
    Malmquist bias at a significance higher than 5 $\sigma$. The favored
    underlying stretch model is a bimodal one, composed of a high-stretch mode
    shared by both young and old environments, and a low-stretch mode that is
    exclusive to old environments. The precise effect of the redshift evolution
    of the intrinsic properties of a SN Ia population on cosmology remains to be
    studied. The astrophysical drift of the SN stretch distribution does affect
    current Malmquist bias corrections, however, and thereby the distances that
    are derived based on SN that are affected by observational selection
effects. We highlight that this bias will increase with surveys covering
increasingly larger redshift ranges, which is particularly important for the
Large Synoptic Survey Telescope.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
enabled the discovery that the Universe's expansion accelerates
\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today
for understanding the properties of dark energy (DE) because they are the only
tool that can precisely map the recent expansion rate ($z<0.5$) when DE is
driving it \citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the
key to directly measuring the Hubble constant ($H_0$) if their absolute
magnitude can be calibrated \citep{riess2016, freedman2019}. Interestingly, the
value of $H_0$ that is derived when the SNe~Ia are anchored to Cepheids
\citep[the Supernovae, $H_0$, for the Equation of State of dark energy
project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
from cosmic microwave background (CMB) data measured by Planck assuming the
standard $\Lambda$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
luminosity is anchored at intermediate redshift by the baryon acoustic
oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
branch technique in place of the Cepheids seems to favor an intermediate value
of $H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
lensing also appear to favor high $H_0$ values \citep{wong2019}.

@@

To model the evolution of the full SN stretch distribution as a function of
redshift, we need to model the SN stretch distribution for each age subsample
given our aforementioned model of the evolution of the fraction of younger and
older SNe~Ia with cosmic time. \cite{rigault2020} presented the relation between
SN stretch and LsSFR measurement, a progenitor age tracer, using the SNfactory
sample. This relation is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory
SNe used in the current analysis. Given the structure of the stretch-LsSFR
scatter plot, our model of the underlying SN~Ia stretch distribution is defined
as follows: the stretch distribution of the younger population
($\log(\mathrm{LsSFR})\geq-10.82$) is modeled as a single normal distribution
$\mathcal{N}(\mu_1, \sigma_1{}^2)$, and the stretch distribution of the older
population ($\log(\mathrm{LsSFR})<-10.82$) is modeled as a bimodal Gaussian
mixture $a\times \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2{}^2)$, where one mode is the same as for the young
population, $a$ representing the relative effect of the two modes.

@@

\begin{table*}
    \centering
    \caption{Best-fit values of the parameters for the base stretch distribution
    model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
569 SN~Ia sample, or the conservative sample (422).}
    \label{tab:modelresults}
    \begin{tabular}{lccccc}
        \hline\hline
        Sample & $\mu_1$ & $\sigma_1$
               & $\mu_2$ & $\sigma_2$
               & $a$ \\
        \hline
        SNfactory & $ 0.41 \pm 0.08$ & $0.55 \pm 0.06$
                  & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$
                  & $ 0.48 \pm 0.08$ \\
        Fiducial & $ 0.37 \pm 0.05$ & $0.61 \pm 0.04$
                 & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$
                 & $ 0.51 \pm 0.09$ \\
        Conservative & $ 0.38 \pm 0.05$ & $0.60 \pm 0.04$
                     & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$
                     & $ 0.47 \pm 0.09$ \\
        \hline
    \end{tabular}
\end{table*}

@@

@@

\bibitem[Rigault et al.(2020)]{rigault2020} Rigault, M., Brinnel, V., Aldering,
G., et al.\ 2020, \aap, 644, A176
